### 2013-09-17 manuscripts/reply/reply.tex
More work on reply.

@@

@@

They also argue that ``\msb can estimate \vmratio{} much better than \numt{}.''
However, the primary objective of \msb is to estimate the posterior probability
of divergence models.
\citet{Oaks2012} demonstrate that even when all assumptions of the model are
met, \vmratio is a terrible model-choice estimator (see plots B, D \& F of
Figure 4), whereas \numt{} performs better.
Furthermore, \vmratio{} is limited to estimating the probability of only a
single divergence model (the one divergence model), and thus its utility for
model choice is extremely limited.

@@

>>>>

\section*{EB}
\citet{Hickerson2013} repeatedly refer the prior distributions used by
\citet{Oaks2012} as ``poorly selected''.

@@

\section*{Our power analysis}

@@

They also argue that ``\msb can estimate \vmratio{} much better than \numt{}.''
However, the primary objective of \msb is to estimate the posterior probability
of divergence models.
\citet{Oaks2012} demonstrate that even when all assumptions of the model are
met, \vmratio is a terrible model-choice estimator (see plots B, D \& F of
Figure 4), whereas \numt{} performs better.
Furthermore, \vmratio{} is limited to estimating the probability of only a
single divergence model (the one divergence model), and thus its utility for
model choice is extremely limited.
The model-choice utility of \vmratio{} is limited to the probability that
this continuous statistic, which can range from zero to infinity, is at its lower
limit of zero. In theory, this point density will always be zero, thus an
arbitrary threshold (0.01 is used throughout the \msb literature) must 
be chosen to make the probability estimable.
However, it is still not surprising to see that it is numerically difficult to
obtain reliable estimate of the probability that the continuous \vmratio{}
statistic is ``near'' its limit of zero.
It is much easier, less subjective, and more interpretable to estimate
the probability of the discrete parameter of the model, \numt{}, is
at its lower limit of one.
Thus, it is not surprising to see this estimator of model probability
perform better.

\section*{General thoughts}
It is not too surprising that the model-averaging approach proposed by
\citet{Hickerson2013} behaves so poorly.
While this approach is trivial to implement, it is not a trivial change
to the the basic \msb model.
To better understand why this is, it might help to step back and get a
sense of the scale of the \msb model.
To do this, we will use the dataset of 22 vertebrate taxon pairs of
\citet{Oaks2012} as an example.

Following the model description and notation of \citet{Oaks2012}, we
will tally up all of the free parameters in the \msb model.

Under the simplest model in \msb (i.e., assuming no migration and
no intra-locus recombination), the number of parameters for each
taxon pair include:
The population sizes of the ancestral and descendant populations
(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
The magnitude of population contraction in each of the descendant
populations (\bottleScalar{1}{} and \bottleScalar{2}{}) and the
timing of these contractions (\bottleTime{}), and the $N-1$ node heights
(coalescent times) of the gene tree that gives rise to the $N$ gene
copies sampled from both populations of the pair.

Because we only have a single locus for each taxon, there are no locus-specific
\myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
\locusRateHetShapeParameter shape parameter for the gamma prior distribution on
\locusMutationRateScalar{}.
Lastly, there are between one and 22 divergence time parameters \divt{} in
the vector \divtvector.
Thus for our Philippines data and under the simplest model in \msb, there are
581--602 free parameters.
Under this rich model, the method is sampling over 1002 divergence models
\citep[i.e., the number of integer partitions of $Y=22$][]{Oaks2012}.

This is a very tall order, and the method only uses four summary statistics
calculated from the sequence alignment of each taxon pair:
$\pi$ \citep{Tajima1983}, $\theta_W$
\citep{Watterson1975}, $\pi_{net}$ \citep{Takahata1985}, and
$SD(\pi-\theta_W)$ \citep{Tajima1989}.
That gives us a total of 88 summarys statistics (four for each of the 22 taxon
pairs), which contain minimal information about many of the $\approx 600$
parameters in the model.
More summary statistics can be used in \msb, but most are highly correlated
with these four statistics (which are even highly correlated among themselves),
and thus contribute little additional information about the data.

This very large number of parameters and divergence models relative to 
summary statistics is undoubtedly the reason the method is so sensitive
to the prior distributions.
It also likely contributes to the methods lack of robustness.
We use robustness here as a measure of a methods insensitivity to model
violations.
Robustness is an extremely important characteristic of a method to gauge its
applicability to empirical data, because we know the model and priors will be
wrong to some degree.

The approach of \citet{Hickerson2013} adds an additional dimension of model
choice to the model. They expand the model to sample over eight prior models.
This extends the original model to having 582-603 free parameters and, more
importantly, sampling over 8016 unique model states across both divergence
models and prior models.
This certainly is a non-trivial extension of the model, and given the methods
very large number of parameters and models relative to the information
used from the data, likely plays a major role in the poor behavior of this
approach.

In theory, the approach of \citet{Hickerson2013} is very appealing.  It sums
over multiple candidate prior models to produce a posterior estimate
marginalized over the uncertainty in prior choice.
In general, Bayesian model-averaging is a powerful approach that leverages a
great strengths of Bayesian statistical procedures, namely the ability to
obtain marginalized estimates that incorporate uncertainty in nuisance
parameters.
However, given the basic \msb model is already struggling to estimate
a huge number of parameters and model probabilities with scant information
from the data, it is not surprising that adding an additional layer of
model choice to the method causes problems.

The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
and biases of \msb are actually similar to those of \citet{Hickerson2013},
but avoid the need for imposing additional model choice.
\citet{Oaks2012} suggest that uniform priors are inappropriate for parameters,
and recommend the use of more cogent probability distributions from the
exponential family.
If we look at the prior distribution on the divergence time parameter \divt{}
imposed by the model-averaging prior approach of \citet{Hickerson2013} we see
it is a mixture of overlapping uniforms with lower limits of zero
(Figure~\ref{figMCTauPrior}).
This looks very much like an exponential distribution.
Thus would be much more natural and easier to simply place a gamma prior (the
exponential being a special case of the gamma) on divergence times.
This would allow the investigator to much better capture their prior
uncertainty in model parameters while avoiding awkward broad uniform
distributions, and without the need of costly model-averaging.
This seems like a much more natural solution to the problems of Lindley's
paradox and sampling error, which are both certainly exacerbated by the fact
that uniform distributions are inappropriately the only choice of prior for
many of the parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
\descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
\bottleScalar{1}{}, \bottleScalar{2}{}, \locusMutationRateScalar,
\migrationRate{}, \recombinationRate;][]{Oaks2012}.
There is no cost whatsoever to this strategy.
Given that ABC methods merely have to draw random values from prior
distributions, hence there are no difficult proposal probability ratios to
calculate, etc., there is no excuse for not using the most appropriate
distributions on parameters with conjugacy in mind.
Also, this approach reduces the ``need'' for empirically guided priors.
The investigator can place the majority of the prior density in regions
of parameters space the believe, a priori, are most plausible, but still
capture uncertainty in the tails of distributions with low density.
As discussed in \citet{Oaks2012} the use of inflexible and unforgiving uniform
priors necessitated the need to use broad priors to avoid excluding the truth a
priori.
Thus, to avoid the   behavior of the method under these priors, cheating
and using the data was a very tempting alternative.

@@

%% Model-choice tau prior
\mFigure{../../model-choice/tau-prior/tau_prior.pdf}{
    The prior distribution on divergence times imposed by the model-averaging prior.
}{figMCTauPrior}
